\newpage
\section{References}

\begingroup
\renewcommand{\section}[2]{}
\begin{thebibliography}{}
	\bibitem{Legende}
	A.M.Legendre. 
	Nouvelles méthodes pour la détermination des orbites des comètes.
	Libraure pour les Mathématiques.
	(1805).
	
	\bibitem{Gauss}
	C.F. Gauss. 
	Theoria Motus Corporum Coelestium in Sectionibus Conicis Solem Ambientum. 
	Hamburg: Frid. Perthes and I.H. Besser.
	(1809).
	
	\bibitem{Hastie}
	T. Hastie, R. Tibshirani, J. Friedman.
	The Elements of Statistical Learning.
	Springer-Verlag, New York.
	(2009).
	
	\bibitem{USGS}
	United States Geological Survey (USGS).
	https://earthexplorer.usgs.gov/
	
	\bibitem{StatMek}
	Lecture notes in Statistical Physics: Statistical Physics - a second course.
	F. Ravndal, E. G. Flekk\oe y.
	(2014).
	
	\bibitem{Coin}
	Simulation stimulation.
	L. Bastick.
	https://www.sumproduct.com/thought/simulation-stimulation
	
	\bibitem{Berkeley}
	Machine Learning Crash Course: Part 4 - The Bias-Variance Dilemma.
	D. Geng, S. Shih.
	https://ml.berkeley.edu/blog/2017/07/13/tutorial-4/
	(2017).
	
	\bibitem{R2}
	Regression Analysis: How Do I Interpret R-squared and Assess the Goodness-of-Fit?
	http://blog.minitab.com/blog/adventures-in-statistics-2/regression-analysis-how-do-i-interpret-r-squared-and-assess-the-goodness-of-fit
	(2013).
	
	\bibitem{Bootstrap}
	Bootstrap Methods: Another Look at the Jackknife.
	B. Efron. 
	Ann. Statist., Volume 7, Number 1, 1-26.
	(1979).
	
	\bibitem{Lecture}
	Lecture notes FYS-STK4155: Regression Methods.
	https://compphysics.github.io/MachineLearning/doc/pub/Regression/pdf/Regression-minted.pdf
	M. Hjorth-Jensen.
	(2018).
	
	\bibitem{Scikit}
	Scikit-learn: Machine Learning in Python.
	F. Pedregosa, G. Varoquaux, A. Gramfort, V. Michel, B. Thiron, O. Grisel, M. Blondel, P. Prettenhofer, R. Weiss, V. Dubourg, J. Vanderplas, A. Passos, D. Cournapeau, M. Brucher, M. Perrot, E. Duchesnay.
    Journal of Machine Learning Research, Volume 12.
	(2011).
	
\end{thebibliography}
\endgroup